---
title: "MPG consumption analysis betweeen automatic and manual transmissions"
author: "Ridzuan Mohamad"
date: "3/19/2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(ggplot2)
library(datasets)

data("mtcars")
```

## Introduction
The Motor Trend, a magazine about the automobile industry. Looking at a data set of a collection of cars, they are interested in exploring the relationship between a set of variables and miles per gallon (MPG) (outcome). They are particularly interested in the following two questions:

1. Is an automatic or manual transmission better for MPG
2. Quantify the MPG difference between automatic and manual transmissions

## Data Analysis & Exploratory
The dataset consist of 11 variables but we only interested only a few variable such as _mpg, cyl, hp, wt, am and gear_. The detail description for the variable as describe in **Appendex A and B**

```{r}
mtcars_clean <- subset(mtcars, select = c("mpg", "cyl", "hp", "wt", "am", "gear"))
```

Than, calculate the mean for both transmission type. The boxplot graph as shown in **Appendix C**
```{r}
c(automatic = mean(mtcars[mtcars$am==0,]$mpg), 
    manual = mean(mtcars[mtcars$am==1,]$mpg))
```

The result shows that the automatic transmission is better than manual transmission in gas consumption. Let take that as our Null Hypothesis ( $H_0$). Now we test the $H_0$ with T-Test.

```{r}
t.test(mtcars[mtcars$am==0,]$mpg, mtcars[mtcars$am==1,]$mpg)
```

The _p-value_ from the T-Test is less than 0.005 so have to reject the $H_0$

## Regression Model

Let creat 2 models for our case:-

1. model.simple which is mpg as outcome and am as predictor.
2. model.full which is mpg as outcome and other variables as predictor

```{r}
model.simple <- lm(mpg ~ am, data = mtcars_clean)
model.all <- lm(mpg ~ ., data = mtcars_clean)
```

```{r}
anova(model.simple, model.all)
```

## Summary

# Appendix

### Appendix A
Variable | Description
---------|----------------------
mpg      | Miles/(US) gallon
cyl	     | Number of cylinders
hp	     | Gross horsepower
wt	     | Weight (1000 lbs)
am	     | Transmission (0 = automatic, 1 = manual
gear	 | Number of forward gears

### Appendix B
```{r}
str(mtcars)
```

### Appendix C
```{r}
with(mtcars, boxplot(mpg ~ am, xlab="Transmission", ylab="MPG",
        main="MPG vs. Transmission Type"))
```

### Appendix D
```{r}
summary(model.simple)
summary(model.all)
```

```{r}
ggplot(mtcars_clean, aes(y=mpg, x=factor(am, labels = c("automatic", "manual")), fill=factor(am)))+
        geom_violin(colour="black")+
        xlab("transmission") + ylab("MPG")
```