---
title: "MPG consumption analysis betweeen automatic and manual transmissions"
author: "Ridzuan Mohamad"
date: "3/19/2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(ggplot2)
library(datasets)

data("mtcars")
```

## Introduction
The Motor Trend, a magazine about the automobile industry. Looking at a data set of a collection of cars, they are interested in exploring the relationship between a set of variables and miles per gallon (MPG) (outcome). They are particularly interested in the following two questions:

1. Is an automatic or manual transmission better for MPG
2. Quantify the MPG difference between automatic and manual transmissions

## Data Analysis & Exploratory
The dataset consist of `r dim(mtcars)[2]` and `r dim(mtcars)[1]` observation data. The detail for each variable as per Appendix A.
```{r}
str(mtcars)
```

Than, calculate the mean for both transmission type and plot the boxplot to visualized the pattern for both transmission types. _(refer Appendix B)_
```{r}
c(automatic = mean(mtcars[mtcars$am==0,]$mpg), manual = mean(mtcars[mtcars$am==1,]$mpg))
```

The result shows that the automatic transmission is better than manual transmission in gas consumption. Let take that as our Null Hypothesis ( $H_0$). Now we test the $H_0$ with T-Test and the probability of Type 1 error is ($\alpha = 0.05$)

```{r}
hypo.result <- t.test(mtcars[mtcars$am==0,]$mpg, mtcars[mtcars$am==1,]$mpg)
hypo.result$p.value
```

However the hypothesis testing show that the _p-value_ < $\alpha$ which force us to reject the $H_0$.

## Regression Model

Firstly, let look at the linear model that corespond to our $H_0$._(Appendix D)_ 

```{r}
model.0 <- lm(mpg ~ am, mtcars)
result.0 <- summary(model.0)
```

The Adjusted R-squared = `r result.0$adj.r.squared`. It shows that the model only able to explain 34% of the variance. 

So, we have to include other variable to our model by using R step() function.

```{r}
best.model <- step(lm(mpg ~ ., data = mtcars), trace=0)
result.1 <- summary(best.model)
result.1
```

The Adjusted R-squared = `r result.1$adj.r.squared` shows that 83% of the variance of the MPG variable. Which mean the variable _wt, qsec and am_ have a correlation affect on _mpg_.

## Summary

Based on the _best.model_ we created; we can assumed that the automatic transmission is `r coef(result.1)[4]` mpg better than manual transmission.

\newpage

# Appendix

### Appendix A - Selected Variable and Description

No.| Variable | Description
---|----------|----------------------
1. |mpg      | Miles/(US) gallon
2. |cyl	     | Number of cylinders
3. |disp	 | Displacement (cu.in.)
4. |hp	     | Gross horsepower
5. |drat	 | Rear axle ratio
6. |wt	     | Weight (1000 lbs)
7. |qsec	 | 1/4 mile time
8. |vs	     | V/S
9. |am	     | Transmission (0 = automatic, 1 = manual
10. |gear	 | Number of forward gears
11. |carb	 | Number of carburetors

### Appendix B - Boxplot MPG vs. Transmission Type
```{r}
ggplot(mtcars, aes(x = factor(am, labels=c("automatic", "manual")), y = mpg, 
                   fill=factor(am, labels=c("automatic", "manual")))) +
    geom_boxplot() +
    scale_x_discrete(name = "Transmission") +
    scale_y_continuous(name = "MPG - Miles per Gallon") +
    ggtitle("MPG vs. Transmission Type") +
    theme(legend.title=element_blank())
```

### Appendix D - Model's Summary 
```{r}
summary(lm(mpg ~ am, mtcars))
summary(best.model)
```
